Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-08-30 paper/paper.tex
Prop 2, Cor 1, Defns, Proofs, Cor env
Notational fixes.

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}

@@

% Maths operators
\newcommand{\nds}{\operatorname{nds}}
\newcommand{\paths}{\operatorname{paths}}
\newcommand{\lessnonind}[1]{\prescript{}{#1}{\rightarrowtail}\ }
\newcommand{\lessind}[1]{\prescript{}{#1}{\hookrightarrow}\ }
\newcommand{\lessmap}[1]{\prescript{}{#1}{\mapsto}\ }

% to do comments
\usepackage{color}
\newcommand{\todo}[1]{{\color{red} {[}{#1}{]}}}

@@

A common induced subgraph of graphs $G$ and $H$ is a graph $P$, together with two induced subgraph
isomorphisms to $G$ and $H$; a maximum common induced subgraph is a common induced subgraph with as
many vertices as possible. An induced $k$-less subgraph isomorphism $P \lessind{k} T$ is equivalent
to a common induced subgraph of $P$ and $T$ with $|\operatorname{V}(P)| - k$ vertices.

If defined similarly, a maximum common non-induced subgraph would allow us to select every vertex in
the smaller of the two graphs, and none of the edges. It is therefore traditional to change the
objective to maximise the number of edges selected, rather than vertices, when a non-induced common
subgraph is sought---this problem is usually called the maximum common \emph{partial} subgraph
problem instead. However, this is not what we will be discussing in this paper: maximum common
subgraph problems are symmetric in their inputs, but when discussing the non-induced case we are
allowing extra edges only in the target graph, not in the pattern.

?? Draw an example.

\subsection{Constraint Models and Algorithms}

@@

\begin{proposition}
    \label{prop:deg}
    Let $p$ be a vertex in $P$ and $t$ a vertex in $T$. For both non-induced and induced
    $k$-less-subgraph isomorphisms, if $p \lessmap t$ then
    $\deg(p) - k \le \deg(t)$.
\end{proposition}
\begin{proof}
Let $p$ be a vertex in $P$, and $t$ a vertex in $T$, with $p\mapsto t$. Then by the definition of subgraph isomorphisms, $\deg(p) \le \deg(t)$. Let $P'$ be $P$ less $k$ vertices and $p$ a vertex in $P'$. Then
\[
\deg_{P}(p)-k \le \deg_{P'}(p) \le \deg_{P}(p) \le \deg(t).
\]
\end{proof}

A vertex $q$ which is adjacent to $p$ is called its \emph{neighbour}.  Let $N(p) = \{ q \in V_p | p
\text{ and } q \text{ are adjacent }\}$ be the set of neighbours of $p$.  The \emph{neighbourhood
degree sequence} (NDS) of $p$, $\nds(p)$, is the (non-ascending) sequence of degrees of its
neighbours.

Let $S = ( s_1 , \ldots , s_n )$ and $T = ( t_1 , \ldots , t_m)$ be two sequences.  We say that $S
\preceq T$ if $n \leq m$ and $\forall s_i \in S$ there exists a distinct $t_j \in T$ with $s_i \leq
t_j$.  (When considering a $k$-less-subgraph isomorphism) we say that $S \prescript{}{k}{\preceq}
T$ if $n - k \leq m$ and $\forall s_i \in S \setminus S_k$, where $S_k$ is a subsequence of $S$
containing up to $k$ members of $S$, there exists a distinct $t_j \in T$ with $s_i - k \leq t_j$.

\begin{proposition}
If $p \prescript{}{k}{\mapsto} t$, then $\nds(p) \prescript{}{k}{\preceq} \nds(t)$.
\end{proposition}

\begin{proof}
Let $p \prescript{}{k}{\mapsto} t$.  Then $\deg(p) - k \leq \deg(q)$, by Proposition 1, which
implies that $|\nds(p)| -k \leq |\nds(t)|$.

Also, $p \prescript{}{k}{\mapsto} t$ implies that for each $q \in N(p)\setminus P_k$, where $P_k$
is some subset of $P$ with $|P_k| \leq k$, $q \prescript{}{k}{\mapsto} u$, where $u \in N(t)$ and
each $u$ is distinct.  Then $\deg(q) - k \leq \deg(u)$, by Proposition 1.  Hence $\nds(p)
\prescript{}{k}{\preceq} \nds(t)$.
\end{proof}

@@

\begin{proposition}
    Let $p,q \in V(P)$ and $t,u\in V(T)$. If $p \lessmap{k} t$ and $q \lessmap{k} u$ then
     $\paths(p, q, 2) - k \le \paths(t, u, 2)$.
\end{proposition}
\begin{proof}
Let $P_{2}$ be the set of all paths of length two between $p$ and $q$,
\begin{align*}
P_{2} = & \{((p,x),(x,q)) : (p,x),(x,q)\in E(P) \} \\
 = & \{((p,x),(x,q)) : x\in N(p) \text{ and } x\in N(q) \} \\
 = & \{((p,x),(x,q)) : x\in N(p)\cap N(q) \} .
\end{align*}
So $\paths(p,q,2) = | N(p)\cap N(q)| \leq deg(p)$.

@@

\begin{proposition}
    For a graph $G$, let $G^{n, \ell}$ be the graph with vertex set $\operatorname{V}(G)$. The
    vertices $p$ and $q$ in $G^{n, \ell}$ are adjacent, if there are at least $n$ simple paths of
    length exactly $\ell$ between $p$ and $q$ in $G$. Then any $k$-less-subgraph isomorphism
    $P \lessnonind{k} T$ induces a new $k$-less-subgraph isomorphism
    $P^{n, 2}\lessnonind{k} T^{n - k, 2}$.
\end{proposition}

>>>>

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}

@@

% Maths operators
\newcommand{\nds}{\operatorname{nds}}
\newcommand{\paths}{\operatorname{paths}}
\newcommand{\lessnonind}[1]{\prescript{}{#1}{\rightarrowtail}\ }
\newcommand{\lessind}[1]{\prescript{}{#1}{\hookrightarrow}\ }
\newcommand{\lessmap}[1]{\prescript{}{#1}{\mapsto}\ }
\newcommand{\lesspreceq}[1]{\prescript{}{#1}{\preceq}\ }
\newcommand{\V}{\operatorname{V}}
% to do comments
\usepackage{color}
\newcommand{\todo}[1]{{\color{red} {[}{#1}{]}}}

@@

A common induced subgraph of graphs $G$ and $H$ is a graph $P$, together with two induced subgraph
isomorphisms to $G$ and $H$; a maximum common induced subgraph is a common induced subgraph with as
many vertices as possible. An induced $k$-less subgraph isomorphism $P \lessind{k} T$ is equivalent
to a common induced subgraph of $P$ and $T$ with $|\V(P)| - k$ vertices.

If defined similarly, a maximum common non-induced subgraph would allow us to select every vertex in
the smaller of the two graphs, and none of the edges. It is therefore traditional to change the
objective to maximise the number of edges selected, rather than vertices, when a non-induced common
subgraph is sought---this problem is usually called the maximum common \emph{partial} subgraph
problem instead. However, this is not what we will be discussing in this paper: maximum common
subgraph problems are symmetric in their inputs, but when discussing the non-induced case we are
allowing extra edges only in the target graph, not in the pattern.
\todo{is this example graph still needed??}
\subsection{Constraint Models and Algorithms}

@@

\begin{proposition}
    \label{prop:deg}
    Let $p$ be a vertex in $P$ and $t$ a vertex in $T$. For both non-induced and induced
    $k$-less-subgraph isomorphisms, if $p \lessmap{k} t$ then
    $\deg(p) - k \le \deg(t)$.
\end{proposition}
\begin{proof}
Let $p$ be a vertex in $P$, and $t$ a vertex in $T$, with $p\mapsto t$. Then by the definition of subgraph isomorphisms, $\deg(p) \le \deg(t)$. Let $P'$ be $P$ less $k$ vertices and $p$ a vertex in $P'$. Then
\[
\deg_{P}(p)-k \le \deg_{P'}(p) \le \deg_{P}(p) \le \deg(t).
\]
\end{proof}

A vertex $q$ which is adjacent to $p$ is called its \emph{neighbour}.  Let $N(p) = \{ q \in \V(P) :
 p \text{ and } q \text{ are adjacent }\}$ be the set of neighbours of $p$.  The \emph{neighbourhood
degree sequence} (NDS) of $p$, $\nds(p)$, is the (non-ascending) sequence of degrees of its
neighbours.

Let $S = ( s_1 , \ldots , s_n )$ and $T = ( t_1 , \ldots , t_m)$ be two sequences.  We say that $S
\preceq T$ if $n \leq m$ and $\forall s_i \in S$ there exists a distinct $t_j \in T$ with $s_i \leq
t_j$.  (When considering a $k$-less-subgraph isomorphism) we say that $S \lesspreceq{k}
T$ if $n - k \leq m$ and $\forall s_i \in S \setminus S_k$, where $S_k$ is a subsequence of $S$
containing up to $k$ members of $S$, there exists a distinct $t_j \in T$ with $s_i - k \leq t_j$.

\begin{proposition}
If $p \lessmap{k} t$, then $\nds(p) \lesspreceq{k} \nds(t)$.
\end{proposition}

\begin{proof}
Let $p \lessmap{k} t$.  Then $\deg(p) - k \leq \deg(q)$, by Proposition 1, which
implies that $|\nds(p)| -k \leq |\nds(t)|$.

Also, $p \lessmap{k} t$ implies that for each $q \in N(p)\setminus P_k$, where $P_k$
is some subset of $P$ with $|P_k| \leq k$, $q \lessmap{k} u$, where $u \in N(t)$ and
each $u$ is distinct.  Then $\deg(q) - k \leq \deg(u)$, by Proposition 1.  Hence $\nds(p)
\lesspreceq{k} \nds(t)$.
\end{proof}

@@

\begin{proposition}
    Let $p,q \in \V(P)$ and $t,u\in \V(T)$. If $p \lessmap{k} t$ and $q \lessmap{k} u$ then
     $\paths(p, q, 2) - k \le \paths(t, u, 2)$.
\end{proposition}
\begin{proof}
Let $P_{2}$ be the set of all paths of length two between $p$ and $q$,
\begin{align*}
P_{2} = & \{((p,x),(x,q)) : (p,x),(x,q)\in E(P) \} \\
 = & \{((p,x),(x,q)) : x\in N(p) \text{ and } x\in N(q) \} \\
 = & \{((p,x),(x,q)) : x\in N(p)\cap N(q) \} .
\end{align*}
So $\paths(p,q,2) = | N(p)\cap N(q)| \leq deg(p)$.

@@

\begin{proposition}
    For a graph $G$, let $G^{n, \ell}$ be the graph with vertex set $\V(G)$. The
    vertices $p$ and $q$ in $G^{n, \ell}$ are adjacent, if there are at least $n$ simple paths of
    length exactly $\ell$ between $p$ and $q$ in $G$. Then any $k$-less-subgraph isomorphism
    $P \lessnonind{k} T$ induces a new $k$-less-subgraph isomorphism
    $P^{n, 2}\lessnonind{k} T^{n - k, 2}$.
\end{proposition}
